---
title: "Retail Sales Forecasting" 
subtitle: "Retail sales Forecasting with Meta's Prophet" 
author: Yuchen Wang 
output: 
  html_document:
    toc: true
    toc_float: true
    theme: spacelab 
    highlight: tango
---

```{r, echo=FALSE}
# This code will display the QMUL logo at the top right of the page
# Do not change this code
htmltools::img(src = knitr::image_uri("images/QMlogo.png"),
               alt = 'logo',
               style = 'position:absolute; top:0; right:0; padding:10px; width:20%;')
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='hide'}
Sys.setlocale("LC_TIME", "English")
```

# Introduction
This project forecasts future retail sales in the US by Meta's Prophet.


# Data
Read monthly retail sales data between 1 January 2016 and 1 January 2026 and convert it into a dataframe with time column *ds* and data column *y*
```{r, warning=FALSE, message=FALSE}
library(prophet)

data_retailsales <- read.csv("data/RSAFS.csv")

prophet_data_retailsales <- data.frame(
  ds = as.Date(data_retailsales$observation_date),
  y = as.numeric(data_retailsales$RSAFS)
)
```

# Time series plot
Plot the diagram of retail sales between 1 January 2016 and 1 January 2026.
```{r}
plot(prophet_data_retailsales$ds, prophet_data_retailsales$y, type = "l",
     main = "Retail Sales Time Series",
     xlab = "Date", ylab = "Retail Sales")
```


The series shows an overall upward trend over time.

# Forecast
Build a prophet model and forecast the future data for the next 12 months
```{r, warning=FALSE, message=FALSE}
model_retailsales <- prophet(prophet_data_retailsales)

future_retailsales <- make_future_dataframe(
  model_retailsales,
  periods = 12,
  freq = "month"
)

forecast_retailsales <- predict(model_retailsales, future_retailsales)
```

# Forecast plot
Plot the diagram of forecast retail sales
```{r}
plot(model_retailsales, forecast_retailsales)
```


The plot shows the forecast for the next 12 months and the uncertainty interval in blue. It suggests that retail sales continue to increase over time.

# Forecast components
Plot the estimated trend and seasonal pattern
```{r, warning=FALSE, message=FALSE}
prophet_plot_components(model_retailsales, forecast_retailsales)
```


The first plot shows a long run trend that retail sales generally incease over time. 
The second plot shows a seasonal pattern that repeats each year.



# Reference
Meta's Prophet Documentary: <https://facebook.github.io/prophet/docs/quick_start.html#r-api>
Retail Sales dataset: <https://fred.stlouisfed.org/series/RSAFS>



